---
title: "P8105: Homework 6"
author: "Molly Martorella"
date: "11/21/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(broom)

```

# Problem 1

Load the data:

```{r}

birth <- read_csv(file = "data/birthweight.csv") %>% 
  mutate(babysex = recode_factor(babysex, 
                                 '1' = "male", 
                                 '2' = "female"),
         frace = recode_factor(frace,
                               '1' = "white",
                               '2' = "black",
                               '3' = "asian",
                               '4' = "puerto_rican",
                               '8' = "other"),
         malform = recode_factor(malform, 
                                 '0' = "absent", 
                                 '1' = "present"),
         mrace = recode_factor(mrace, 
                               '1' = "white",
                               '2' = "black",
                               '3' = "asian",
                               '4' = "puerto_rican",
                               '8' = "other"))

```

Checked for missing data and appropriate variable types using `skimr::skim(birth)`, which showed no missing data and numeric variables were recoded to factors where appropriate. The dataset contains `r nrow(birth)` observations of `r ncol(birth)` variables, which describe characteristics of the mother pre and postpartum, the father, and the newborn baby.

## Generating a linear regression model

```{r}

molly_model = lm(data = birth, bwt ~ smoken + ppbmi + momage + fincome)



```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.

- smoking --> small, obese - diabetes --> large baby
- Need to account for malformations -- maybe exclude that data.
- How to account for gestational age?
- does mother's height or bmi have an effect?
- parity could have an effect - smaller 1st babies versus later?
- how to account for previous number of low birth weight or small for gestational age babies?

## Comparison of model to provided models

Create models to predict birthweight:
1. `basic_baby` - length at birth and GA as predictors
2. `all_baby` - head circumference, length, sex, and all interactions

```{r}

basic_baby = lm(data = birth, bwt ~ blength + gaweeks)

basic_baby %>% 
  tidy() %>% 
  knitr::kable(digits = 3)

```

Both baby length and gestational age have a significant rela on baby weight that is significant. 

```{r}
all_baby = lm(data = birth, bwt ~ bhead + blength + babysex + bhead*blength + bhead*babysex + blength*babysex + bhead*blength*babysex)

all_baby %>% 
  tidy() %>% 
  knitr::kable(digits = 3)

```

All variables and interactions between them have a significant effect on final baby weight except for the interaction between head circumference and baby length.

